---
output: github_document
---

Load the necessary packages and read in the data.csv file into R (use an R chunk)

```{r}
# loading common packages, not sure if they'll be used
library(tidyverse)
library(ggplot2)

# see if R can read the file
read.csv("data/data.csv")

# asign data to an object
data <- read.csv("data/data.csv")

# view data
view(data)
```

Manipulate the data frame as necessary so that you can calculate average duration, f0 and intensity as a function of lexical stress (extra points if you can create a plot)

```{r}
# data tidy

# one variable per cell
# separate word & stressed condition (1, stressed; 2, unstressed)
# change durationV column name

data_tidy <- data |>
  separate(
    col = info,
    into = c("word", "stress_cond"),
    sep = "_") |>
  rename(duration = durationV)

# data is ready!
```

Calculate average duration as a function of lexical stress

```{r}
data_duration <- data_tidy |> 
  select(word, stress_cond, duration)

aggregate(duration ~ stress_cond, FUN = mean, data = data_duration)
```

Calculate average f0 as a function of lexical stress

```{r}
data_f0 <- data_tidy |> 
  select(word, stress_cond, f0)

aggregate(f0 ~ stress_cond, FUN = mean, data = data_f0)
```

Calculate average intensity as a function of lexical stress

```{r}
# create data frame exclusive for dur
data_int <- data_tidy |> 
  select(word, stress_cond, int)

aggregate(int ~ stress_cond, FUN = mean, data = data_int)
```

Extra points if you can create a plot

```{r}
# create plot for average duration as a function of lexical stress

ggplot(
  data = data_duration,
  mapping = aes(x = stress_cond, y = duration, color = word)
) +
  geom_point(aes(color = word)) +
  labs(
    title = "Average duration as a function of lexical stress",
    x = "Stress Condition: 1, stressed; 2, unstressed", y = "Vowel Duration (ms)",
    color = "Word")

# create plot for average f0 as a function of lexical stress

ggplot(
  data = data_f0,
  mapping = aes(x = stress_cond, y = f0, color = word)
) +
  geom_point(aes(color = word)) +
  labs(
    title = "Average f0 as a function of lexical stress",
    x = "Stress Condition: 1, stressed; 2, unstressed", y = "Vowel F0",
    color = "Word")

# create plot for average intensity as a function of lexical stress

ggplot(
  data = data_int,
  mapping = aes(x = stress_cond, y = int, color = word)
) +
  geom_point(aes(color = word)) +
  labs(
    title = "Average intensity as a function of lexical stress",
    x = "Stress Condition: 1, stressed; 2, unstressed", y = "Vowel Intensity",
    color = "Word")
```


